---
title: "GDP Per Capita"
author: "Cornelius Tanui"
date: "7/19/2021"
output:
  html_document:
    code_folding: hide
    theme: united
    highlight: tango

---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      include = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
```

Collect garbage and disable graphics, and clear environment :-- 

```{r Clear Evrironment}
gc()
graphics.off()
rm(list = ls(all.names = TRUE))
```

Load required libraries 

```{r Load Packages}
# load Packages
library(tidyverse)   # for data wrangling 
library(flexdashboard)  # framework for outputs
library(shiny)       # for responsive plots
library(plotly)      # for interactive plots 
library(DT)          # for data tables
library(ggplot2)     # for graphing 
library(ggpmisc)     # for helper functions of ggplot2
library(gapminder)   # for source dataset
library(tidymodels)  # for modelling  
library(vip)         # for variable importance
library(kableExtra)  # for html tables
library(gganimate)   # for animating a plot
library(gifski)       # for gif outputs from animated plots
library(av)         # for video output
```

## Plots {.tabset .tabset-pills}

### Static

This is a basic static `ggplot2` plot of `lifeExp ~ gdpPercap` by `pop` and `continent`. 

```{r plot1}
p <- gapminder %>%
  filter(year==1977) %>%
  ggplot( aes(x = gdpPercap, 
              y = lifeExp, 
              size = pop, 
              color = continent)) +
  geom_point() +
  geom_jitter() + 
  geom_smooth(method = "lm",
              formula = y ~ log(x),
              color = "red",
              se = FALSE)+
  labs(title = "Life Expectancy vs GDP Per Capita",
       y = "Life Expectancy",
       x = "GDP Per Capita")+
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = c(0.88, 0.4),
        legend.title = element_blank(),
        legend.background = element_blank())

ggsave(filename = "Life Expectancy vs GDP Per Capita.png",
       width = 7, height = 7,
       units = "in")

ggplotly(p)
```


### Animated

This is a basic animated `ggplot2` plot of `lifeExp ~ gdpPercap` by `pop` and `continent`. 

```{r plot2}
p1 <- gapminder %>%
  ggplot( aes(x = gdpPercap, 
              y = lifeExp, 
              size = pop, 
              color = continent)) +
  geom_point() +
  geom_jitter() + 
  geom_smooth(method = "lm",
              formula = y ~ log(x),
              color = "red",
              se = FALSE)+
  labs(title = "Life Expectancy vs GDP Per Capita",
       y = "Life Expectancy",
       x = "GDP Per Capita")+
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = c(0.88, 0.4),
        legend.title = element_blank(),
        legend.background = element_blank())+
  transition_time(year) +
  ease_aes('linear')


p1
```

set the default engine

```{r default model}
lm_model <- linear_reg() %>% 
  set_engine('lm') %>% # adds lm implementation of linear regression
  set_mode('regression')
```

Get the `gapminder` dataset

```{r data}
data <- gapminder
```

Fit model

```{r fit}
lm_fit <- lm_model %>% 
  fit(lifeExp ~ gdpPercap + pop + continent, data = data)
```

View `fit` properties

```{r table}
tidy(lm_fit) %>%
  filter(p.value < 0.05) %>%
  kbl(caption = "Significant Factors") %>%
  kable_classic(full_width = F, html_font = "Cambria")
```
  
View $R^2$ 

```{r R Square}
glance(lm_fit)%>%
  kbl(caption = "R-Squared Value") %>%
  kable_classic(full_width = F, html_font = "Cambria")
```

variable importance

```{r importance}
vip(lm_fit, 
    aesthetics = list(fill = 2:7))
```
