---
title: "GDP Per Capita"
author: "Cornelius Tanui"
date: "7/19/2021"
output:
  html_document:
    code_folding: hide
    theme: united
    highlight: tango

---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      include = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
```

Collect garbage and disable graphics, and clear environment :-- 

```{r Clear Evrironment}
gc()
graphics.off()
rm(list = ls(all.names = TRUE))
```

Load required libraries 

```{r Load Packages}
# load Packages
library(tidyverse)   # for data wrangling 
library(flexdashboard)  # framework for outputs
library(shiny)       # for responsive plots
library(plotly)      # for interactive plots 
library(DT)          # for data tables
library(ggplot2)     # for graphing 
library(ggpmisc)     # for helper functions of ggplot2
library(gapminder)   # for source dataset
library(tidymodels)  # for modelling  
library(vip)         # for variable importance
library(kableExtra)  # for html tables
library(gganimate)   # for animating a plot
library(gifski)       # for gif outputs from animated plots
library(av)         # for video output
```

## Plots {.tabset}

### Static

This is a basic static `ggplot2` plot of `lifeExp ~ gdpPercap` by `pop` and `continent`. 

```{r plot1}
p <- gapminder %>%
  filter(year==1977) %>%
  ggplot( aes(x = gdpPercap, 
              y = lifeExp, 
              size = pop, 
              color = country)) +
  geom_point() +
  geom_jitter() + 
  geom_smooth(method = "lm",
              formula = y ~ log10(x), 
              se = FALSE, 
              size = 0.5, 
              color = "black", 
              linetype="dotted")+
  scale_colour_manual(values = country_colors) +
  labs(title = "Life Expectancy vs GDP Per Capita",
       y = "Life Expectancy",
       x = "GDP Per Capita")+
  scale_x_continuous(labels = dollar) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = c(0.88, 0.4),
        legend.title = element_blank(),
        legend.background = element_blank())

ggsave(filename = "Life Expectancy vs GDP Per Capita.png",
       width = 7, height = 7,
       units = "in")

ggplotly(p)
```


### GG Animate

This is a basic [`gganimate`](https://gganimate.com/) animation using `ggplot2`, a plot of `lifeExp ~ gdpPercap` by `pop` and `continent`. 

```{r plot2}
p2 <- gapminder %>%
  ggplot(aes(x = gdpPercap, 
             y = lifeExp, 
           size = pop, 
           colour = country)) +
  geom_point(alpha = 0.7, 
             show.legend = FALSE) +
  geom_smooth(method = "lm",
              formula = y ~ log10(x), 
              se = FALSE, 
              size = 0.5, 
              color = "black", 
              linetype="dotted")+
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  labs(title = "Life Expectancy vs GDP Per Capita by 'Year: {frame_time}'",
       y = "Life Expectancy",
       x = "Logarithmic GDP Per Capita") +
  theme(plot.title = element_text(hjust = 0.5))+
  transition_time(year) +
  ease_aes('linear')

```

Save as gif (.gif)

```{r gif}
# save as gif
p2_gif <- animate(p2, 
                  fps = 10, 
                  width = 750, 
                  height = 450,
                  renderer = gifski_renderer())

anim_save(p2_gif, filename = "life expectancy.gif")

p2_gif
```

Save as video (.mp4)

```{r video}
# save as a video
p2_vid <- animate(p2,
                  width = 800,
                  height = 450,
                  renderer = ffmpeg_renderer())

anim_save(p2_vid, filename = "life expectancy.mp4")

p2_vid

```


### Plotly Animate

This is a basic [plotly](https://plotly.com/) animation using and `ggplot2` plot of `lifeExp ~ gdpPercap` by `pop` and `continent`. 

```{r plot3}
p3 <- gapminder %>%
  ggplot(aes(gdpPercap, lifeExp, color = continent)) +
  geom_point(aes(size = pop, frame = year, ids = country)) +
  scale_x_log10() +
   scale_colour_manual(values = continent_colors) +
  labs(title = "Life Expectancy vs GDP Per Capita by Year",
       y = "Life Expectancy",
       x = "Logarithmic GDP Per Capita") +
  theme(plot.title = element_text(hjust = 0.5)) 

ggplotly(p3)%>% 
  animation_opts(1000, 
                 easing = "elastic", 
                 redraw = FALSE) %>% 
  animation_button(x = 1, 
                   xanchor = "right",
                   y = 0,
                   yanchor = "bottom") %>%
  animation_slider(currentvalue = list(prefix = "YEAR ", 
                                       font = list(color="blue")))

```

set the default engine

```{r default model}
lm_model <- linear_reg() %>% 
  set_engine('lm') %>% # adds lm implementation of linear regression
  set_mode('regression')
```

Get the `gapminder` dataset

```{r data}
data <- gapminder
```

Fit model

```{r fit}
lm_fit <- lm_model %>% 
  fit(lifeExp ~ gdpPercap + pop + continent, data = data)
```

View `fit` properties

```{r table}
tidy(lm_fit) %>%
  filter(p.value < 0.05) %>%
  kbl(caption = "Significant Factors") %>%
  kable_classic(full_width = F, html_font = "Cambria")
```
  
View $R^2$ 

```{r R Square}
glance(lm_fit)%>%
  kbl(caption = "R-Squared Value") %>%
  kable_classic(full_width = F, html_font = "Cambria")
```

variable importance

```{r importance}
ggplotly(vip(lm_fit, 
    aesthetics = list(fill = 2:7)), tooltip = "Importance")
```






